---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("dce78b11de43007eaba2a37f3d725ebcfc920857")
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
acs5 <- load_variables(2019, "acs5", cache = TRUE)
View(acs5)

```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
#Write code here to read the RDS file
removed_ppp_loans<-read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first.
**A1** Baltimore City has the lowest median age. The age is 35.4.  

```{r}
md <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2019)

md <- md %>% arrange(estimate) %>% top_n(24, estimate)

```

**Q2** Which Maryland counties have a median age of at least 50?
**A2** Worcester County and Talbot County. Their estimates are 50.4 and 50.5 respectively. 

```{r}
md %>%
  filter(estimate>=50)
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? Start with the code you wrote for Q4 and add a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** The counties with the highest percentages of 5-17 year olds that speak Spanish are: Prince George's County, Montgomery County, Caroline County. 

```{r}
md_span_at_home<- get_acs(geography = "county",
              variables = c(speak_span = "B16007_004"),
              state = "MD",
              year = 2019) %>%
      rename(speak_span=estimate)

md_youth_pop<- get_acs(geography = "county",
              variables = c(pop = "B06001_003"),
              state = "MD",
              year = 2019) %>%
      rename(youth_pop=estimate)

md_youth_pop %>%
  inner_join(md_span_at_home, by=c("GEOID", "NAME")) %>%
  select(-variable.x, -variable.y) %>%
   mutate(percent = (speak_span/youth_pop)*100) %>%
  arrange(desc(percent))
```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4** The counties that stand out to me are Montgomery County and Prince George's County because their populations that speak Spanish are already pretty close. MoCo's margin of error is really large and if that was added to their population that speaks Spanish, then it would be greater than PG County's population that speaks Spanish. 
  The top five counties don't really have much overlap with their margin of errors besides MoCo and PG. The difference is not significant for MoCo and PG, but for the rest of them the difference is significant. Although the populations for each of the top five counties are a wide range, the percentages of each of the populations start getting more consistent and closer together as you move more into Maryland and further from DC. 
  Garrett County is the furthest from DC and closest to West Virginia, so if PG and MoCo had a large percentage of Spanish speaking youth and they're close to DC, then it would make sense that Garrett's population is very low. Also, the county is extremely small, which may also explain why that percentage is low.  

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** The 5 digit zip code that had the most removed PPP applications is 60620. The percentage of black population in that zip code is 95.6934619.

```{r}
#total population by zip code
total_pop <- get_acs(geography = "zcta",
              variables = c(population = "B01003_001"),
              year = 2019) %>%
    rename(total_population=estimate)
#total black population by zip code
black_pop <- get_acs(geography = "zcta",
              variables = c(population = "B02001_003"),
              year = 2019)%>%
    rename(total_black_population=estimate)
#joining total pop and black pop
total_with_black<-total_pop%>%
  inner_join(black_pop, by=c("GEOID"))%>%
  select(-variable.x, -variable.y)%>%
  mutate(percent_black_pop = (total_black_population/total_population)*100)
#5 digit ppp loans
removed_ppp_loans <- removed_ppp_loans %>%
  mutate(zip5 = str_sub(zip, 1, 5))

#joining ppp with population
all_data_together<-removed_ppp_loans%>%
  inner_join(total_with_black, by=c("zip5"="GEOID"))

#chart
all_data_together%>%
  group_by(zip5,percent_black_pop)%>%
  summarise (count_loans = n()) %>%
  arrange(desc(count_loans))
```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed application in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6** I would tell the editor that 8/10 of the zip codes with the most removed applications have a majority of Black people with almost all of them over 75%. Also all of these zip codes except one are in Cook County, Illinois, which has a very obvious systemic problem of oppression of Black people. The only zip code that seems different is Atlanta, Georgia. It has a low population of Black people and it's the only one not in Chicago. So, my thoughts for that are maybe there's a larger class divide in Atlanta with more lower class people, and the rejected applications are from people in the lower class.

**Q7** What other columns from the PPP data could be useful in reporting on what happened to this removed loans?
**A7** Other columns from the PPP that could be useful are maybe gender and business type. I think then you could see more trends, which will make it easier to investigate.
