---
title: "Reverse Engineering Project"
author: Jenna Cohen, Taneen Momeni, Max Boyarsky and Jamie Pinzon
date: 10/9/2021
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been fatally shot by police since 2015] (https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/ )

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
#turn off scientific notation
options(scipen=999)
#load tidyverse
library(tidyverse)
#load lubridate 
library(lubridate)
#load refinr
library(refinr)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

fatal_police_shootings<- read_csv("data/fatal-police-shootings-data.csv")

# Clean required data and prepare for analysis if needed. 

#counts rows to get total number of fatal shootings
total_fatal_shootings <- fatal_police_shootings %>% 
  nrow()

#filters and counts to get number of Black women fatally shot
total_black_women_shot <- fatal_police_shootings %>% filter(gender == "F" & race == "B") %>% nrow()

#filters and counts to get number of fatal shootings where race is known
race_known <- fatal_police_shootings %>% filter(!is.na(race)) %>% nrow() 




```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [“Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths.”]
* **Analysis summary**: [I was able to confirm the finding for the first part of the sentence, which stated that black women accounted for less than 1 percent of overall fatal shootings. However, when I calculated the remaining two parts of the sentence, I couldn't confirm the finding. The sentence states that black women account for 20% of women shot, but I got 18%. The more troubling of the two not confirmed was the last part. My work found that 12.6% of all unarmed deaths were black women, but the sentence states they account for 28%. A possible explanation would be that they included the "undetermined" entries in the unnamed category as well. Another possible explanation is because the story was published in Sept of 2020, and the data base was most recently updated in Sept. 2021. The trends in the data are relatively similar, but I believe that's why I cannot replicate their findings exactly.]


```{r}
# Put code to reverse engineer sentence here
#(part 1)Black women have accounted for less than 1 percent of the overall fatal shootings

total_black_women_shot_percentage = (total_black_women_shot/race_known)*100
total_black_women_shot_percentage

# Display results of code below this codeblock
```

```{r}
# Put code to reverse engineer sentence here

# (part 2) Black women, who are 13 percent of the female population, account for 20 percent of the women shot
total_women_shot = fatal_police_shootings %>% 
   #total women shot
  filter(gender == "F") %>% nrow()
  #divide black women by total women shot
  black_women_out_of_women_shot = (total_black_women_shot/total_women_shot)*100
# Display results of code below this codeblock
   black_women_out_of_women_shot
```

```{r}
# Put code to reverse engineer sentence here

# (part 3) Black women account for 28 percent of the unarmed deaths.
total_unarmed = fatal_police_shootings %>% 
   #total women shot
  filter(armed == "unarmed") %>% nrow()
  #divide black women by total unarmed
  black_women_out_of_unarmed = (total_black_women_shot/total_unarmed)*100
# Display results of code below this codeblock
   black_women_out_of_unarmed 
```

### Sentence 2

* **Sentence text**: [“About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.”]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] My entire analysis is a bit off from the data listed in the sentence. I believe that is because the story was published in Sept of 2020, and the database was most recently updated in Sept. 2021. The trends in the data are relatively similar, but I believe that's why I cannot replicate their findings exactly. 

```{r}
# Put code to reverse engineer sentence here

#create new df of women shot
women_df <- fatal_police_shootings %>% 
  filter(gender=="F")
#count rows
  nrow(women_df)
  #294 women
#filter for signs_of_mental_illness
women_mental_illness<-  women_df %>% 
  filter(signs_of_mental_illness=="TRUE")
#count rows 
  nrow(women_mental_illness)
  #95 women with mental illness
#divide women by signs of mental illness and *100
(95/294)*100
  #32.3% 
    
#create new df of men shot
men_df <- fatal_police_shootings %>% 
  filter(gender=="M")
#count rows
  nrow(men_df)
  # 6303 men
#filter for signs_of_mental_illness
men_mental_illness<-  men_df %>% 
  filter(signs_of_mental_illness=="TRUE")
#count rows 
  nrow(men_mental_illness)
  #1399 men with mental illness
#divide men by signs of mental illness and *100
  (1399/6303)*100
  #22.2%
# Display results of code below this codeblock
```


### Sentence 3

* **Sentence text**: [“As with fatal police shootings of men, the vast majority of the women killed were armed with a potential weapon at the time, although slightly less often: 89 percent of the women were armed, compared to 91 percent of the men.”]
* **Analysis summary**: [I think I was able to confirm the findings. It was unclear what the Washington Post meant by "undetermined" in the armed column. If "undetermined" meant that the weapon on the person was undetermined (and not that it meant it was undetermined if the person shot had a weapon), then the percentages that came out of that were close to what the Post said. So, the Post most likely thought that undetermined meant armed. When considering "undetermined" as being unarmed, then the percentages are off by a few percentage points.] 

```{r}
# Put code to reverse engineer sentence here
#Each gender is run twice- one including the undetermined number count as armed and one counting undetermined as unarmed. 

#89 percent of women were armed (with undetermined considered armed)
fatal_police_shootings%>%
  filter(gender=="F" & armed != "unarmed")%>%
  summarise(
    count = n(),
    percent_women_armed= (count/total_women_shot)*100
  ) 
#89 percent of women were armed (with undetermined considered unarmed)
fatal_police_shootings%>%
  filter(gender=="F" & armed != "unarmed" & armed != "undetermined")%>%
  summarise(
    count = n(),
    percent_women_armed_without_undet= (count/total_women_shot)*100
  ) 
#filters and counts men fatally shot
  total_men_shot <- fatal_police_shootings %>% filter(gender == "M") %>% nrow()
#91 percent of men were armed (with undetermined considered armed)
fatal_police_shootings%>%
  filter(gender=="M" & armed != "unarmed")%>%
  summarise(
    count = n(),
    percent_men_armed= (count/total_men_shot)*100
  ) 
#91 percent of men were armed (with undetermined considered unarmed)
fatal_police_shootings%>%
  filter(gender=="M" & armed != "unarmed" & armed != "undetermined")%>%
  summarise(
    count = n(),
    percent_men_armed_without_undet= (count/total_men_shot)*100
  ) 

```




### Sentence 4

* **Sentence text**: ["Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. While about twice as many White women were shot and killed as Black women, White women account for five times the Black female population."]
* **Analysis summary**: [For the first part, my result was that police killed 28 unarmed women, not 26. The second part I found that there were 15 White women and 5 Hispanic women instead, hence 2 more people in the outcome. The differences in the result might be because the story was published in September 2020, but the data that was used was updated in September 2021, which might have added new information.]

```{r}
# Put code to reverse engineer sentence here
#Since 2015, police have killed 26 unarmed women.
fatal_police_shootings%>%
  filter(gender=="F", armed=="unarmed")

#14 were White, seven were Black, four were Hispanic and one’s race was unknown
white_unarmed_women<-fatal_police_shootings%>%
  filter(gender=="F", armed=="unarmed", race=="W")%>%
  summarise(count = n())

black_unarmed_women<-fatal_police_shootings%>%
  filter(gender=="F", armed=="unarmed", race=="B")%>%
  summarise(count = n())

hispanic_unarmed_women<-fatal_police_shootings%>%
  filter(gender=="F", armed=="unarmed", race=="H")%>%
  summarise(count = n())
  
# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: ["By race, 147 of the women killed were White, 48 Black and 29 Hispanic. Five were Native American, four were Asian and three were other races. In 11 cases, race could not be determined."]

* **Analysis summary**: [I filtered for gender (F), grouping by race, and summarising by the number of women killed using the summarise function with the “n count.” It showed that 175 women killed were white, 53 were black, 5 were Native American, and 5 were asian. 20 cases could not be determined. This does not confirm the finding because most of the numbers are greater than the Post's numbers, which is most likely because the dataset was updated in 2021, and the Post's story was published in 2020.]

```{r}
# Put code to reverse engineer sentence here
fatal_police_shootings%>%
  filter(gender== "F" ) %>% 
  group_by(race) %>% 
  summarise(
    women_killed = n())
```


-30-

