---
title: "lab_11"
author: "sean mussenden"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. You'll need tidyverse, tidycensus, janitor, sf and tigris, and you'll also need to install your census API key.

```{r}
library(tidyverse)
library(sf)
library(janitor)
library(tigris)
library(tidycensus)
```

## Load data

**Task** Create a codeblock and load the WV slice of the PPP loan data, which is in the data folder and is called lab_11.rds.

```{r}
wv_ppp_loans<-read_rds("data/lab_11.rds")
```


## Questions 

**Q1.** You are interested in doing some reporting on how West Virginia companies in the "Mining, Quarrying, and Oil and Gas Extraction" industry (https://www.naics.com/six-digit-naics/?code=21) used PPP loans during the pandemic (West Virginia is a big coal mining state).  
Then produce a choropleth map that shows West Virginia counties shaded according to the total amount of loans to that industry per 100,000 population. 

Use the magma color palette, and shade on a "log" scale. Also produce a table that shows the total amount of loans per 100,000 population for each county, ranked from highest to lowest, along with the total amount of loans and the total number of loans. 

```{r}
wv_mining<-wv_ppp_loans%>%
  mutate(naics_code = str_sub(naics_code, start=1L, end=2L))%>%
  filter(naics_code == 21)%>%
  group_by(project_county_name)%>%
  summarise(total_amount = sum (amount), total_loans=n())%>%
  rename(county=project_county_name)
#population for WV
wv_county_population <- get_acs(geography = "county", 
              variables = c(population = "B01001_001"), 
              state = "WV",
              geometry=TRUE)%>%
              rename(county=NAME)%>%
              mutate(county = toupper(str_remove_all(county,", West Virginia|County"))) %>%
              mutate(county = str_trim(county,side="both"))%>%
              rename(population=estimate)%>%
              select(county, population,geometry)

#joining
wv_loans_per_100k<-wv_county_population%>%
  left_join(wv_mining)%>%
  mutate(total_amount=replace_na(total_amount,0))%>%
  mutate(total_loans=replace_na(total_loans,0))%>%
  mutate(amount_per_100k=total_amount/population*100000)%>%
  arrange(desc(amount_per_100k))

#da map
ggplot() + 
  geom_sf(data=wv_loans_per_100k, aes(fill=amount_per_100k)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma", trans="log")
```

Then answer these questions:

* Based on the map, what region of the state appears to have gotten more loans in this industry? What geographic pattern do you notice in the data?
* Based on the table, what county got the highest total loan amount? What was the total amount per 100K population, the total amount of loans (not population adjusted), how many loans were there, and what's the population?

**A1.** I think the middle of West Virginia (like the north and south middle areas) has gotten more loans in this industry. The geographic pattern is that the closer the counties are to the western and eastern boarders of WV, the less loans they got in this industry. The county with the highest total loan amount was Ritchie county. The total amount per 100k: 77752742.79, total amount of loans: 7653980.00, number of loans: 9, and the population: 9844. 

**Q2.** You want to understand WHY the spatial pattern you saw above exists.  So you grab a shapefile with the latitude and longitude coordinates of every mine permit in West Virginia from the state Department of Environmental Protection. 

Here's the url: https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json. 

(It's in a slightly different format than the .shp file in the lab. It's another type of shapefile called "geojson".  It will work just the same, though, despite that difference. You can read it directly into R without manually downloading it, by putting the url in the st_read() function we learned about in this week's prelab.) 

This data includes both currently operating mines and old ones that have closed.  You want to examine only mines that have a permit expiration date after Sept. 1, 2021. 

Make a map with a point for each one of these mine permits and put a blank map with the outline of West Virginia counties (which you can get from Tigris) behind it. 

```{r}
#loading in the url
wv_mine_permits<-st_read("https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json")

#permits after 9/1/21
wv_new_permits<-wv_mine_permits%>%
  filter(expiredate>="2021/1/9")%>%
  filter(expiredate!="No Date")

#dots
wv_new_permits %>%
  ggplot() + 
  geom_sf() + 
  theme_minimal()

#tigris
counties<-counties()
wv_counties<-counties%>%
  filter(STATEFP==54)

#map
wv_counties %>%
  ggplot() + 
   geom_sf(data=wv_counties)+
  geom_sf(data=wv_new_permits) +
  theme_minimal()

```

What spatial pattern do you see? How does it compare to the spatial pattern in the first map you made?

**A2.** There is a high concentration of mines in the southern and northern regions of West Virginia. This pattern is similar to the first map because the last map had a high concentration of loan amounts per 100k in the southern and middle northern regions of the state.

**Q3.** To confirm the spatial comparison you identified in question 2, let's create a new map that layers the mine location points on top of our PPP choropleth map we created in question 1. What do you notice about the amount of mines in the county that has the highest total loan amount per 100K?

```{r}
#layering maps
ggplot() + 
  geom_sf(data=wv_loans_per_100k, aes(fill=amount_per_100k)) + 
  geom_sf(data=wv_new_permits) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma", trans="log")
```

**A3.** There is barely one mine in the county that has the highest total loan amount per 100k. There are more mines in counties that have lower total loan amounts per 100k.

**Q4.** Starting from the original WV ppp dataframe, examine the loans to companies in "Mining, Quarrying, and Oil and Gas Extraction" in the county with the highest total loan amount per 100K (you can use the table you created in question 1 to identify the top county)?
* What single company accounts for the biggest chunk of the total loan amount? 
* What specific industry are they in, based on their 6-digit NAICS code? Look them up on the web. What services do they provide?
* How many jobs did they save, according to the data?
* What's your hypothesis for why a company in their specific industry would need to seek a PPP loan during the pandemic?
```{r}
#ritchie county is the top county
ritchie_mines<-wv_ppp_loans%>% 
  filter(project_county_name=="RITCHIE")%>%
  mutate(naics_code = str_detect(naics_code, "^21"))

ritchie_summed<-ritchie_mines%>% 
  group_by(name)%>%
  summarise(total_amount=sum(amount))%>%
  arrange(desc(total_amount))
```
**A4.** The company that amounts for the biggest chunk of the total loan amount is Hall Drilling, LLC. Their naics code is 213111, and the specific industry that they're in is Drilling Oil and Gas Wells. They saved 387 jobs. My hypothesis for them needing to seek a PPP loan during the pandemic is that because of mandates and needing to stay home, not many people were driving often and needed gas. Because there was less need for gas, this company may not have had a lot of business and therefore needed to apply for a loan.  
-30-
